---
title: "Data Analysis"
author: "Hao Wang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Merged Data and Clean it

```{r, echo=FALSE}
library(MASS)
library(dplyr)
library(ggplot2)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/vdem-merge.csv?token=AX7YEY61G5zquz7jnDfL0Sl5k-3Bw8upks5Y9BGxwA%3D%3D")

```


```{r, echo=FALSE, include=FALSE}
names(mydata)
summary(mydata) #tons of missing data
```


# Get ready for the dependent variables:

I measure the budget stability as the simple euclidean distance of the between-year percentage differences. It can be written in the following equation: $S_t$ is the stability index of a certain year $t$. Since government budget has various categories: $p_{it}$ denotes the percentage of $i$th category of total expenditure. $S_t$ will increase as the difference between $p_{it}$ and $p_{it-1}$ increases. 

$$S_t = \sqrt{\sum_{i = 1}^i (p_{it} -p_{it-1})^2} ~~~~~~~~~~~~~~~~~~~~ (1)$$ 

```{r generate DV, echo=FALSE, include=FALSE}


```

