---
title: "Data Analysis"
author: "Hao Wang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Merging several sources

## Working Environment 

```{r packages, message=FALSE, warning=FALSE}
library(foreign)
library(dplyr)
library(readstata13)
library(countrycode)
setwd("D:/Dropbox/01Research Project/Budget-Vdem")
#getwd()
```


## Merge Data

```{r, warning=FALSE}
GSRE <- read.dta13('GSRE_extended_1_0.dta')
VDEM <- read.dta13('V-Dem-DS-CY+Others-v6.2.dta')
colnames(VDEM)[10] <- "cowcode"
#This keeps all points from GSRE
mydata <-  merge(GSRE, VDEM, by=c("cowcode","year"), all.x = TRUE) 
```


# Data summary

## For V-dem project
V-dem file is download from [stata-and-other]()


### Related to democracy
```{r}
#Deliberative democracy index v2x-delibdem
#To what extend is deliberatuve democracy achieved? aggregated index
summary(mydata$v2x_delibdem)

#give reasoned justification of their positions on public policy
summary(mydata$v2dlreason)
#give reasoned justification of their positions on common good
summary(mydata$v2dlcommon)
#give respect for counterarguments
summary(mydata$v2dlcountr)
#range of consultation
summary(mydata$v2dlconslt)
#engaged society
summary(mydata$v2dlengage)
#-----------------------------------------------------------------

#equal democracy index
#equal protection of individual rights
summary(mydata$v2xeg_eqprotec)
#equal distribution of resource 
summary(mydata$v2xeg_eqdr)

#constraints on executive branches
#judicial constraints
summary(mydata$v2x_jucon)
#legislative constraints
summary(mydata$v2xlg_legcon)

#civil society organization participation index
#rountely consulted by policy maker
summary(mydata$v2cscnsult)

#---------------------------------------------------
#party system institutionalization
summary(mydata$v2xps_party)
```


### Related to corruption

```{r}
#political corruption index-aggregated
summary(mydata$v2x_corr)
#public sector corruption index
summary(mydata$v2x_pubcorr)
#executure corruption index
summary(mydata$v2x_execorr)
#-------------------------------------------

#election vote buying 
summary(mydata$v2elvotbuy)
#voting irregularities
summary(mydata$v2elirreg)

```




### Related to Federalism and Decentralization 






It seems weird, Jones et. al (2017) paper looks at federalism, and parliamentalism 






## GSRE Part

### Percentage spent as of total expenditure
expend_security_EXP expenddefence_EXP exp_public_order_EXP wagessalaries_EXP subpentrans_EXP pensions_EXP total_welfare_EXP education_EXP health_EXP social_protection_EXP housing_EXP owelfarespend_EXP



### Percentage spent as of GDP
expend_security_GDPGSRE expenddefence_GDPGSRE exp_public_order_GDPGSRE wagessalaries_GDPGSRE subpentrans_GDPGSRE pensions_GDPGSRE total_welfare_GDPGSRE education_GDPGSRE health_GDPGSRE social_protection_GDPGSRE housing_GDPGSRE owelfarespend_GDPGSRE

