---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
    template: D:/Dropbox/Materials/svm-r-markdown-templates/svm-latex-ms.tex
title: "Veto Player and Fiscal Policy Stability"
thanks: "Replication files are available on the author's Github account (http://github.com/haowang666). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
author:
- name: Hao Wang
  affiliation: Arizona State University
abstract: "Veto player theory [@Tsebelis2002] predicts that the number of veto players influencing policy stabilities. While studies in OECD countries have shown supportive evidence [@TsebelisChang2004], there is few work on policy stability in nondemocracies. This project uses a new dataset from GSRE (Global State Revenues and Expenditures dataset) and perform an empirical test on veto player and budget stability in authoritarian countries. Preliminary analysis show that even in authoritarian countries, institutional constrians (veto players) lead to incremental budget changes. "
keywords: "veto player, public policy"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: double
bibliography: D:/Dropbox/Bibliography-database/Haowang.bib
biblio-style: apsr
---






Punctuated Equilibrium Theory (PET) argues that government budget shifts over and under attention to certain policy areas lead to long periods of stability and short periods of radical changes. Most empirical evidence, however, is drawn from developed democracies. Following Baumgartner et al. (2015) and Lam and Chan (2015), we explore the determinants of policy stability in different authoritarian regimes. We extend the existing theory by examining the variations among authoritarian countries. Our results suggest that institutionalization in the policy making process is an important factor that explains cross national variation.




# Argument 
Veto player $\Rightarrow$ unable to change policy rapidly $\Rightarrow$ long term incremental changes and short-term rapid changes $\Rightarrow$ punctual equilibrium [@DerekBaumgartner2016].





# Data
Data in this project comes from various sources. 

GSRE: 


```{r load data, message=FALSE, tidy=TRUE, warning=FALSE, echo=FALSE}
library(readstata13)
library(MASS)
library(ggplot2)
library(plm)
library(Zelig)
library(MASS)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/vdem-merge.csv?token=AX7YEY61G5zquz7jnDfL0Sl5k-3Bw8upks5Y9BGxwA%3D%3D")
```





```{r, echo=FALSE, include=FALSE}
names(mydata)
summary(mydata) #tons of missing data
```


# Get ready for the dependent variables:

I measure the budget stability as the simple euclidean distance of the between-year percentage differences. It can be written in the following equation: $S_t$ is the stability index of a certain year $t$. Since government budget has various categories: $p_{it}$ denotes the percentage of $i$th category of total expenditure. $S_t$ will increase as the difference between $p_{it}$ and $p_{it-1}$ increases. 

$$S_t = \sqrt{\sum_{i = 1}^i (p_{it} -p_{it-1})^2} ~~~~~~~~~~~~~~~~~~~~ (1)$$ 



```{r generate DV, echo=FALSE, include=FALSE}
#generate DV


```














<!--The folowing are some examples 

```{r coefplot1, eval=FALSE, echo=FALSE, cache=FALSE, tidy=TRUE}
suppressMessages(library(Zelig))
suppressMessages(library(arm))
suppressMessages(library(coefplot))

data(voteincome)

voteincome$z.age <- arm::rescale(voteincome$age)
voteincome$z.education <- arm::rescale(voteincome$education)
voteincome$z.income <- arm::rescale(voteincome$income)

M1 <- glm(vote ~ z.age + female + z.education + z.income,
            data=voteincome, family=binomial)

coefplot(M1)
```


```{r, echo=FALSE, eval=TRUE, tidy = TRUE, cache=FALSE, fig.cap="A Coefficient Plot"}
suppressMessages(library(Zelig))
suppressMessages(library(arm))


data(voteincome)

voteincome$z.age <- arm::rescale(voteincome$age)
voteincome$z.education <- arm::rescale(voteincome$education)
voteincome$z.income <- arm::rescale(voteincome$income)

M1 <- glm(vote ~ z.age + female + z.education + z.income,
            data=voteincome, family=binomial)

arm::coefplot(M1)
```

To get `knitr` to present the results of a table, add `results="asis"` to the brackets to start the R code chunk. The ensuing output will look like this (though the table may come on the next page). 

```{r, echo=FALSE, eval=TRUE, tidy = TRUE, size="small", cache=FALSE, results="asis"}
suppressMessages(library(Zelig))
suppressMessages(library(stargazer))
suppressMessages(library(arm))

data(voteincome)

voteincome$z.age <- arm::rescale(voteincome$age)
voteincome$z.education <- arm::rescale(voteincome$education)
voteincome$z.income <- arm::rescale(voteincome$income)


M1 <- glm(vote ~ z.age + female + z.education + z.income,
            data=voteincome, family=binomial)

stargazer(M1, title="A Handsome Table", header=FALSE)
```

-->


<!-- This is for footnote: Adding `echo="FALSE"` inside the brackets to start the R chunk will omit the presentation of the R commands. It will just present the table. This provides substantial opportunity for authors in doing their analyses. Now, the analysis and presentation in the form of a polished manuscript can be effectively simultaneous.[^4]

[^4]: I'm not sure if I'm ready to commit to this myself since my workflow is still largely derived from [Rob J. Hyndman's example](http://robjhyndman.com/hyndsight/workflow-in-r/). However, *knitr* has endless potential, especially when analyses can stored in cache, saved as chunks, or loaded in the preamble of a document to reference later in the manuscript.

# Add reference
Inline refer: @Nichter2008, in bracket [@Nichter2008], suppress[Gonzales et al. -@Gonzalez-Ocantos2014] -->



<!--
# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent
-->
