---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Dropbox/Materials/svm-r-markdown-templates/svm-latex-ms.tex
title: "Veto Player and Fiscal Policy Stability"
thanks: "Replication files are available on the author's Github account (http://github.com/haowang666). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
author:
- name: Hao Wang
  affiliation: Arizona State University
abstract: "Veto player theory [@Tsebelis2002] predicts that the number of veto players influencing policy stabilities. While studies in OECD countries have shown supportive evidence [@TsebelisChang2004], there is few work on policy stability in nondemocracies. This project uses a new dataset from GSRE (Global State Revenues and Expenditures dataset) and perform an empirical test on veto player and budget stability in authoritarian countries. Preliminary analysis show that even in authoritarian countries, institutional constrians (veto players) lead to incremental budget changes. "
keywords: "veto player, public policy"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: single
bibliography: ~/Dropbox/Bibliography-database/Haowang.bib
biblio-style: apsr
---





# Intro
Veto player theorys [@Tsebelis2002] defines 'veto players' as individuals or institutions whose agreement is required for a change of the status quo. This theory predicts that: when the number of veto players increase, the winning set that can defeat status quo will shrink, which in turn leads to higher policy stability. Since veto player is ultimately related to the level of institutional constraints, a collary is that institutional checks leads to more stable, incremental policy outcomes. With many checks and balances in the government, it will be harder to move policies from status quo equibilibrium. 

@TsebelisChang2004 apply veto player theory in the budget changes of the 19 OECD countries. In their analysis, parties with more polarized positions are modeled as potential veto players who could have blocked the policy proposals. 



Punctuated Equilibrium Theory (PET) argues that government budget shifts over and under attention to certain policy areas lead to long periods of stability and short periods of radical changes. Most empirical evidence, however, is drawn from developed democracies. Following Baumgartner et al. (2015) and Lam and Chan (2015), we explore the determinants of policy stability in different authoritarian regimes. We extend the existing theory by examining the variations among authoritarian countries. Our results suggest that institutionalization in the policy making process is an important factor that explains cross national variation.




# Argument 

## Institutional Constraints

Veto player $\Rightarrow$ unable to change policy rapidly $\Rightarrow$ long term incremental changes and short-term rapid changes $\Rightarrow$ punctual equilibrium [@DerekBaumgartner2016].


[@Henisz2000]

[@KeeferStasavage2003]

[@JohBevan2012]


## Policy Punctuations





# Data
Data in this project comes from various sources. The dependent variable comes from the GSRE project (Global State Revenues and Expenditures dataset). GSRE is a comprehensive budget dataset based on the previous released historical documents from the International Monetary Fund (IMF). Comparing with the IMF [COFOG]() dataset, GSRE increases coverage and accuracy of budgeting data for most authoritarian regimes and some democratic regimes. Since GSRE is built on IMF historical documents, it covers all independent states that have been or are the members of the IMF and are being coded as an authoritian regime in the [@Geddesetal2014] dataset. 

Data on deliberative democracies and other regime-related variables come from the Variaties of Democracy (Vdem) project [@Vdem2016]. Unlike the widely used democracy index like Polity [@Polity2015], Vdem provides multidimensional meassurements of regimes, including both democracies and autocracies. 


Data on institutional constraints come from the political constraints index [@Henisz2000]. Henisz develops a measure of institutional commitment that is objective, extensive, and based in positive political theory. He uses a quantitative model to capture the competitiveness portion of the definition of democracy (competitiveness and participation) with a proxy of number of independent veto points over policy outcomes and distribution of preferences of those actors. The measure is objective, with clear rules for measurement and aggregation, however it incorporates some of the Polity data coding for “independent judiciary” that gives it a slight subjective bias. POLCON is based on strong assumptions about each actor’s veto power. The measures are strongly correlated with the ICRG indexes and the Polity “Executive Constraint” index. “The new release of the political constraint dataset expands the scope of coverage to as many as 234 countries over the period 1800–2001. It also corrects a small number of computational, coding and factual errors in the previous release. The new database also includes country codes from the Cross-national time series data archive, Polity and the World Bank to facilitate matching this dataset to other international datasets that you may have. Finally, it contains the component variables used to construct the political constraint indexes.”




Data on decentralization draws from the Political Institution Index [@Becketal2001]. 



```{r load data, message=FALSE, tidy=TRUE, warning=FALSE, echo=FALSE}
library(MASS)
library(ggplot2)
library(plm)
library(Zelig)
library(dplyr)
library(tidyr)
library(stargazer)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/vdem-merge.csv?token=AX7YEY61G5zquz7jnDfL0Sl5k-3Bw8upks5Y9BGxwA%3D%3D")
#code polity -88 -66 -77 to na
mydata <-
  mydata %>%
  mutate(e_democ = replace(e_democ, e_democ == -88, NA), e_autoc = replace(e_autoc, e_autoc == -88, NA)) %>%
  mutate(e_democ = replace(e_democ, e_democ == -77, NA), e_autoc = replace(e_autoc, e_autoc == -77, NA)) %>%
  mutate(e_democ = replace(e_democ, e_democ == -66, NA), e_autoc = replace(e_autoc, e_autoc == -66, NA)) 
mydata <- select(mydata, -subpentrans_EXP, -subpentrans_GDPGSRE, -X)
```





```{r, echo=FALSE, include=FALSE}
names(mydata)
#summary(mydata) #tons of missing data
```


# Measuring Dependent Variables 

I measure the budget stability as the simple euclidean distance of the between-year percentage differences. It can be written in the following equation: $S_t$ is the stability index of a certain year $t$. Since government budget has various categories: $p_{it}$ denotes the percentage of $i$th category of total expenditure. $S_t$ will increase as the difference between $p_{it}$ and $p_{it-1}$ increases. 

$$S_t = \sqrt{\sum_{i = 1}^i (p_{it} -p_{it-1})^2} ~~~~~~~~~~~~~~~~~~~~ (1)$$ 




\newpage

# Appendix



## Descriptive Statistics


**summary statistics of deliberative democracy**




```{r, echo=FALSE, message=FALSE, tidy=TRUE, results="asis"}
# results ="asis" add to table

delib <- select(mydata, v2x_delibdem: v2dlengage)
stargazer(delib, title= "Deliberative Democracy Statistics", style = "apsr", header = FALSE,
          covariate.labels=c("Deliberative Democracy", "Justification on Public Policy", 
                             "Justification on Common Goods", "Respect for Counterarguments", 
                             "Range of Consultation", "Range of Engagement"))
```

**summary statistics on institutional constriants**

```{r, tidy=TRUE, message=FALSE, echo=FALSE, results="asis"}
const <- select(mydata, v2x_jucon, v2xlg_legcon, v2xps_party, e_democ, e_autoc, e_h_polcon5, e_h_polcon3) 
stargazer(const, title = "Institutional Constraints", header = FALSE,
          covariate.labels = c("Judical Constraints", "Legislative Constraints",
                               "Institutionalization of Party", "Institutionalizaed Democracy",
                               "Institutionalizaed Autocracy", "Political Constraints Index-3", 
                               "Political Constraints Index-5"))
```


**summary statistics on federalism**

```{r, tidy=TRUE, message=FALSE, echo=FALSE, results="asis"}
fed <- select(mydata, v2x_feduni, e_dpi_author)
stargazer(fed, title = "Division of Power (centrl-regional) Index", header = FALSE,
          covariate.labels = c("Division of Power", "State government Authority"))

```


**Dependent Variables**
In the following table I report the components of dependent variables. It is measured as the raw expenditure as of 2005 constant USD. 

```{r DV table, tidy=TRUE, echo=FALSE, results='asis'}
expenditure <- select(mydata, c(78:89))
stargazer(expenditure, title = "Components of Budget Stability Measurements", header = FALSE, digits = 3)
```





The distribution of dependent variable without multiple imputation looks like this 







# Missing Cases
GSRE contains lots of missing cases. To avoid losing statistical powers and potential bias due to listwise deletion, this article employs multiple impuation of the GSRE part data. Results reported in the paper are from the first imputation. Appendix includes the rest 4 imputations. 

The total missing map is showing in the correspoding figure. 

```{r map, message=FALSE, warning=FALSE, echo=FALSE}
library(Amelia)
GSRE <- select(mydata, c(78:89))
missmap(GSRE)
```




I also calculated resutls without multiple imputation: points that are missing in the GSRE dataset is set to be 0. Theoretically in this situation missing cases will contribute zero effects to the policy stability indicator. 



```{r DV generating, echo=FALSE, message=FALSE, warning=FALSE}
#generating dependent variables 
#The first step is to calculate the DV 
#select GSRE
GSRE <- select(mydata, c(78:89))

GSRE$sum <- apply(GSRE, 1, sum, na.rm= TRUE)
GSRE$cowcode <- mydata$cowcode
GSRE$year <- mydata$year

#create percentage variables 
dv1 <- 
  GSRE %>%
  mutate(expend_security_p = expend_security / sum,
         expenddefence_p = expenddefence / sum,
         exp_public_order_p = exp_public_order / sum,
         wagessalaries_p = wagessalaries / sum,
         subpentrans_p = subpentrans / sum,
         pensions_p = pensions / sum,
         total_welfare_p =  total_welfare / sum,
         education_p =  education / sum,
         health_p = health / sum,
         social_protection_p = social_protection / sum,
         housing_p = housing / sum,
         owelfarespend_p = owelfarespend / sum
        )
```



```{r DV generating 2, echo=FALSE, message=FALSE, warning=FALSE}
#create lagged variable
dv1 <- 
    dv1 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.subpentrans_p = lag(subpentrans_p, 1), 
           lag.pensions_p = lag(pensions_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1), 
           lag.owelfarespend_p = lag(owelfarespend_p, 1))


#code missings as 0
dv1[is.na(dv1)] <-0

#calculate stability index
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (subpentrans_p - lag.subpentrans_p)^2 +
      (pensions_p - lag.pensions_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 +
      (owelfarespend_p - lag.owelfarespend_p)^2

#code irregular yb to 1.
yb[yb >1] <- 1 

```


The density plot of stability index without imputation is shown in the following figure

```{r}
mydata$EXP <- sqrt(yb)
d <- ggplot(data = mydata) +
     geom_density(aes(x=EXP)) +
     ggtitle("Density Plot of Policy Stability Index") +
     xlab("Policy Stability Index") 
d
```








<!--The folowing are some examples 

```{r coefplot1, eval=FALSE, echo=FALSE, cache=FALSE, tidy=TRUE}
suppressMessages(library(Zelig))
suppressMessages(library(arm))
suppressMessages(library(coefplot))

data(voteincome)

voteincome$z.age <- arm::rescale(voteincome$age)
voteincome$z.education <- arm::rescale(voteincome$education)
voteincome$z.income <- arm::rescale(voteincome$income)

M1 <- glm(vote ~ z.age + female + z.education + z.income,
            data=voteincome, family=binomial)

coefplot(M1)
```
-->

