---
title: "Data Analysis"
author: "Hao Wang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Merging several sources



## Merge Data

```{r, warning=FALSE}
library(foreign)
library(dplyr)
library(readstata13)

setwd("~/Dropbox/01Research Project/Budget-Vdem")
getwd()
GSRE <- read.dta13("GSRE_extended_1_0.dta")
VDEM <- read.dta13("V-Dem-DS-CY+Others-v6.2.dta")
colnames(VDEM)[10] <- "cowcode"
#This keeps all points from GSRE
mydata <-  merge(GSRE, VDEM, by=c("cowcode","year"), all.x = TRUE) 
```


# Data summary

## For V-dem project
V-dem file is download from [stata-and-other]()


### Related to democracy
```{r}
# v2x_delibdem v2dlreason v2dlcommon v2dlcountr v2dlconslt v2dlengage

#Deliberative democracy index v2x-delibdem
#To what extend is deliberatuve democracy achieved? aggregated index
summary(mydata$v2x_delibdem)

#give reasoned justification of their positions on public policy
summary(mydata$v2dlreason)
#give reasoned justification of their positions on common good
summary(mydata$v2dlcommon)
#give respect for counterarguments
summary(mydata$v2dlcountr)
#range of consultation
summary(mydata$v2dlconslt)
#engaged society
summary(mydata$v2dlengage)
#-----------------------------------------------------------------
# v2xeg_eqprotec v2xeg_eqdr v2x_jucon v2xlg_legcon v2cscnsult
#equal democracy index
#equal protection of individual rights
summary(mydata$v2xeg_eqprotec)
#equal distribution of resource 
summary(mydata$v2xeg_eqdr)

#constraints on executive branches
#judicial constraints
summary(mydata$v2x_jucon)
#legislative constraints
summary(mydata$v2xlg_legcon)

#civil society organization participation index
#rountely consulted by policy maker
summary(mydata$v2cscnsult)

#---------------------------------------------------
#party system institutionalization
summary(mydata$v2xps_party)

#---------------------------------------------------
#legislative constraints 
#legislature questions officials in practice
summary(mydata$v2lgqstexp)
#legislature investigates in practice
summary(mydata$v2lginvstp)
#executive oversight: other body engages in the investigation of the executive branch
summary(mydata$v2lgotovst)

#--------------------------------------------------
#veto legislation by law (Elkins et al. 2012) CCP project
#summary(mydata$e_ccp_leg_in_1)

#--------------------------------------------------
#minimal competitive elections Skaaning et al. 2015
summary(mydata$e_competition)

#-------------------------------------------------
#Marshall and Jaggers 2013
#institutionalizaed democracy
summary(mydata$e_democ)
#institutionalized autocracy
summary(mydata$e_autoc)

#-------------------------------------------------
#institutional constraints as a whole
summary(mydata$e_h_polcon5)
```


### Related to corruption and clientelism

```{r corription}
#political corruption index-aggregated
summary(mydata$v2x_corr)
#public sector corruption index
summary(mydata$v2x_pubcorr)
#executure corruption index
summary(mydata$v2x_execorr)
#-------------------------------------------

#election vote buying 
summary(mydata$v2elvotbuy)
#voting irregularities
summary(mydata$v2elirreg)

#-------------------------------------------
#party linkage: from programmatic to clientelist 
summary(mydata$v2psprlnks)
#executure bribery and corruption exchange 
summary(mydata$v2exbribe)
#executive embezzlement and theft 
summary(mydata$v2exembez)
#public sector corrupt exchanges 
summary(mydata$v2excrptps)

#------------------------------------------
#on delivery of puclic goods and private goods 
#particularistic to puclic funded
summary(mydata$v2dlencmps)


#--------------------------------------------
#Cite: Kaufman et al. 2010
#world bank government efffectiveness
summary(mydata$e_wbgi_gee)

#world bank rule of law index
summary(mydata$e_wbgi_rle)

```


### Political Equality

```{r}
# equal power over socioeconomic conditions
summary(mydata$v2pepwrses)
# equal power over social groups 
summary(mydata$v2pepwrsoc)
```



### Related to Federalism and Decentralization 

```{r decentralization}
#regional government exist 
summary(mydata$v2elreggov)
#regional offices relative power 
summary(mydata$v2elrgpwr)

#local offices exists
summary(mydata$v2ellocgov)
#local offices relative power
summary(mydata$v2ellocpwr)

#division of power index (central-regional-local)
summary(mydata$v2x_feduni)

#Database of political institutions (Beck et al. 2001)
#state government authority over taxation
summary(mydata$e_dpi_author)

```


### Related to information and media

```{r}
#government censorship: media
summary(mydata$v2mecenefm)
#internet censorship
summary(mydata$v2mecenefi)

```


### Demographics (controls)

```{r}
#education: average year of education among citizens older than 15
summary(mydata$e_peaveduc)

#gdp per capita (logged)
summary(mydata$e_migdppcln)

#gdp growth
summary(mydata$e_migdpgro)

#income inequality: UNU-Wider 2008
summary(mydata$e_peginiwi)

#Oil reserves: Haber and Menaldo 2011
# in billions of barrels
summary(mydata$e_reserves_billions)
# Oil production per capita: Haber and Menaldo 2011
summary(mydata$e_Total_Oil_Income_PC)
# Government revenues from oil, gas and minerals
summary(mydata$e_Fiscal_Reliance)

#------------------------------------------------
#Armed conflicts International: Brecke 2001
summary(mydata$e_miinteco)
#Internal armed conflicts 
summary(mydata$e_miinterc)
#civil war: Haber and Menaldo 2011
summary(mydata$e_Civil_War)

```


It seems weird, Jones et. al (2017) paper looks at federalism, and parliamentalism 


## GSRE Part

### Percentage spent as of total expenditure
expend_security_EXP expenddefence_EXP exp_public_order_EXP wagessalaries_EXP subpentrans_EXP pensions_EXP total_welfare_EXP education_EXP health_EXP social_protection_EXP housing_EXP owelfarespend_EXP



### Percentage spent as of GDP
expend_security_GDPGSRE expenddefence_GDPGSRE exp_public_order_GDPGSRE wagessalaries_GDPGSRE subpentrans_GDPGSRE pensions_GDPGSRE total_welfare_GDPGSRE education_GDPGSRE health_GDPGSRE social_protection_GDPGSRE housing_GDPGSRE owelfarespend_GDPGSRE

### Raw expenditure
expend_security expenddefence exp_public_order wagessalaries subpentrans pensions total_welfare education health social_protection housing owelfarespend


## Subset Data

```{r}
trim <- mydata[,c("cowcode","year","country","v2x_delibdem",
                  "v2dlreason", "v2dlcommon", "v2dlcountr",                                                                     "v2dlconslt", "v2dlengage", "v2xeg_eqprotec", 
                  "v2xeg_eqdr", "v2x_jucon", "v2xlg_legcon", 
                  "v2cscnsult", "v2xps_party", "v2lgqstexp",
                  "v2lginvstp", "v2lgotovst", 
                  "e_competition", "e_democ", "e_autoc",
                  "v2x_corr", "v2x_pubcorr", "v2x_execorr",
                  "v2elvotbuy", "v2elirreg", "v2psprlnks",
                  "v2exbribe", "v2exembez", "v2excrptps",
                  "v2dlencmps", "e_wbgi_gee", "e_wbgi_rle",
                  "v2pepwrses", "v2pepwrsoc", "v2elreggov",
                  "v2elrgpwr", "v2ellocgov", "v2ellocpwr",
                  "v2x_feduni", "e_dpi_author", "v2mecenefm",
                  "v2mecenefi", "e_peaveduc", "e_migdppcln",
                  "e_migdpgro", "e_peginiwi", "e_reserves_billions",
                  "e_Total_Oil_Income_PC", "e_Fiscal_Reliance", "e_miinteco",
                  "e_miinterc", "e_Civil_War", "e_h_polcon3", "e_h_polcon5",
                  "expend_security_EXP",  "expenddefence_EXP", "exp_public_order_EXP", 
                  "wagessalaries_EXP","subpentrans_EXP", "pensions_EXP", "total_welfare_EXP",                                   "education_EXP", "health_EXP", "social_protection_EXP", 
                  "housing_EXP", "owelfarespend_EXP", "expend_security_GDPGSRE",                                                "expenddefence_GDPGSRE", "exp_public_order_GDPGSRE", 
                  "wagessalaries_GDPGSRE", "subpentrans_GDPGSRE", 
                  "pensions_GDPGSRE", "total_welfare_GDPGSRE",                                                                  "education_GDPGSRE", "health_GDPGSRE", 
                  "social_protection_GDPGSRE", "housing_GDPGSRE", 
                  "owelfarespend_GDPGSRE" , "expend_security",
                  "expenddefence", "exp_public_order", "wagessalaries",
                  "subpentrans", "pensions", "total_welfare",
                  "education", "health",
                  "social_protection", "housing",
                  "owelfarespend"
                  )]
names(trim)

write.csv(x = trim, file = "vdem-merge.csv")

```




# Clean Some Mis-coded items
```{r}
library(dplyr)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/vdem-merge.csv")
#some cleaning
#code polity -88 -66 -77 to na
mydata <-
  mydata %>%
  mutate(e_democ = replace(e_democ, e_democ == -88, NA), e_autoc = replace(e_autoc, e_autoc == -88, NA)) %>%
  mutate(e_democ = replace(e_democ, e_democ == -77, NA), e_autoc = replace(e_autoc, e_autoc == -77, NA)) %>%
  mutate(e_democ = replace(e_democ, e_democ == -66, NA), e_autoc = replace(e_autoc, e_autoc == -66, NA)) 
mydata <- select(mydata, -subpentrans_EXP, -subpentrans_GDPGSRE, -X)
#get rid of two annoying observations 
newdata <- filter(mydata, wagessalaries_EXP >1)
mydata <- setdiff(mydata, newdata)

write.csv(x = mydata, file = "vdem-merge-clean.csv")


```






```{r}
#generating dependent variables 
#The first step is to calculate the DV 
#select GSRE
GSRE <- select(mydata, c(78:89))
GSRE$sum <- apply(GSRE, 1, sum, na.rm= TRUE)
GSRE$cowcode <- mydata$cowcode
GSRE$year <- mydata$year

#create percentage variables 
dv1 <- 
  GSRE %>%
  mutate(expend_security_p = expend_security / sum,
         expenddefence_p = expenddefence / sum,
         exp_public_order_p = exp_public_order / sum,
         wagessalaries_p = wagessalaries / sum,
         subpentrans_p = subpentrans / sum,
         pensions_p = pensions / sum,
         total_welfare_p =  total_welfare / sum,
         education_p =  education / sum,
         health_p = health / sum,
         social_protection_p = social_protection / sum,
         housing_p = housing / sum,
         owelfarespend_p = owelfarespend / sum
        )

#save data for multiple imputation: percentage one
raw <- select(dv1, cowcode, year, c(16:27))
write.csv(x = raw, file = "raw.csv")

#get original one
raw2 <- GSRE 
write.csv(x = raw2, file = "raw2.csv")

#get exp of total expenditure percentage data
dvp <-  select(mydata, cowcode, year, c(56:66))
write.csv(x = dvp, file = "rawp.csv")

```



# Multiple Imputation : based on raw percentage
```{r rawp, include=FALSE, warning=FALSE}
library(dplyr)
library(Amelia)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/vdem-merge-clean.csv")
rawp <- select(mydata, cowcode, year, c(57:67))
names(rawp)
summary(rawp)
#get rid of pension
rawp <-select(rawp, -pensions_EXP)

set.seed(3)
a.out <- amelia(rawp, ts = "year", cs = "cowcode", m=5)
write.amelia(a.out, format  = "csv", file.stem = "imp")
```


# Adding DV into mydata frame

without multiple imputation

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/vdem-merge-clean.csv")
#This code is using the existing percentage index
dv2 <- select(mydata, c(57:67), cowcode, year)
dv2 <- select(dv2, -pensions_EXP)


#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.wagessalaries_EXP = lag(wagessalaries_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (wagessalaries_EXP - lag.wagessalaries_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_0 <- sqrt(yb)
#recode 0 to missing
mydata$stability_0[mydata$stability_0 ==0] <-NA
```







# Multiple imputation: based on actual expenditure percentage

```{r}
library(dplyr)
imp1 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp1.csv")
imp2 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp2.csv")
imp3 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp3.csv")
imp4 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp4.csv")
imp5 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp5.csv") 

#-----------------------------------------------------------------------------------------------
dv2 <- imp1
#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.wagessalaries_EXP = lag(wagessalaries_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (wagessalaries_EXP - lag.wagessalaries_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_1 <- sqrt(yb)
#recode 0 to missing
mydata$stability_1[mydata$stability_1 ==0] <-NA

#--------------------------------------------------------------------------
dv2 <- imp2
#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.wagessalaries_EXP = lag(wagessalaries_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (wagessalaries_EXP - lag.wagessalaries_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_2 <- sqrt(yb)
#recode 0 to missing
mydata$stability_2[mydata$stability_2 ==0] <-NA

#--------------------------------------------------------------------------
dv2 <- imp3
#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.wagessalaries_EXP = lag(wagessalaries_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (wagessalaries_EXP - lag.wagessalaries_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_3 <- sqrt(yb)
#recode 0 to missing
mydata$stability_3[mydata$stability_3 ==0] <-NA

#--------------------------------------------------------------------------
dv2 <- imp4
#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.wagessalaries_EXP = lag(wagessalaries_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (wagessalaries_EXP - lag.wagessalaries_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_4 <- sqrt(yb)
#recode 0 to missing
mydata$stability_4[mydata$stability_4 ==0] <-NA

#--------------------------------------------------------------------------
dv2 <- imp5
#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.wagessalaries_EXP = lag(wagessalaries_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (wagessalaries_EXP - lag.wagessalaries_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_5 <- sqrt(yb)
#recode 0 to missing
mydata$stability_5[mydata$stability_5 ==0] <-NA

#--------------------------------------------------------------------------
#export to doable file
write.csv(x = mydata, file = "mydata.csv")
```









<!--
#Need to abandon the following codes!!!!

# Multiple Imputation DV : based on percentage
```{r, include=FALSE}
library(dplyr)
library(Amelia)
raw <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/raw.csv")

#keeps hitting warning, getting rid of some variables
raw <- select(raw, -subpentrans_p, -X, -pensions_p, - owelfarespend_p)

set.seed(3)
a.out <- amelia(raw, ts = "year", cs = "cowcode", m=5)

write.amelia(a.out, format  = "csv", file.stem = "impp")
```


## Original results 

```{r, message=FALSE, include=FALSE}
library(dplyr)
library(Amelia)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/vdem-merge.csv?token=AX7YEY61G5zquz7jnDfL0Sl5k-3Bw8upks5Y9BGxwA%3D%3D")
#code polity -88 -66 -77 to na
mydata <-
  mydata %>%
  mutate(e_democ = replace(e_democ, e_democ == -88, NA), e_autoc = replace(e_autoc, e_autoc == -88, NA)) %>%
  mutate(e_democ = replace(e_democ, e_democ == -77, NA), e_autoc = replace(e_autoc, e_autoc == -77, NA)) %>%
  mutate(e_democ = replace(e_democ, e_democ == -66, NA), e_autoc = replace(e_autoc, e_autoc == -66, NA)) 
mydata <- select(mydata, -subpentrans_EXP, -subpentrans_GDPGSRE, -X)

#generating dependent variables 
#The first step is to calculate the DV 
#select GSRE
GSRE <- select(mydata, c(78:89))
GSRE <- select(GSRE, owelfarespend, pensions, summarysubpentrans)


GSRE$sum <- apply(GSRE, 1, sum, na.rm= TRUE)
GSRE$cowcode <- mydata$cowcode
GSRE$year <- mydata$year

#create percentage variables 
dv1 <- 
  GSRE %>%
  mutate(expend_security_p = expend_security / sum,
         expenddefence_p = expenddefence / sum,
         exp_public_order_p = exp_public_order / sum,
         wagessalaries_p = wagessalaries / sum,
         total_welfare_p =  total_welfare / sum,
         education_p =  education / sum,
         health_p = health / sum,
         social_protection_p = social_protection / sum,
         housing_p = housing / sum
        )
#create lagged variable
dv1 <- 
    dv1 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1))


#code missings as 0
dv1[is.na(dv1)] <-0

#calculate stability index
#subpentran, pension and ownwelfare are excluded
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 

#code irregular yb to 1.
yb[yb >1] <- 1 
mydata$stability_0 <- sqrt(yb)


```


## Original DV: based on raw percentage. 





## From first imputation 
```{r, include=FALSE, message=FALSE}
library(dplyr)

impp1 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/impp1.csv")
impp2 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/impp2.csv")
impp3 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/impp3.csv")
impp4 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/impp4.csv")
impp5 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/impp5.csv")  

#create lagged variable
dv1 <- 
    impp1 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1))
#code missings as 0
dv1[is.na(dv1)] <- 0

#calculate stability index
#subpentran, pension and ownwelfare are excluded
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 

#code irregular yb to 1.
yb[yb >1] <- 1 
mydata$stability_1 <- sqrt(yb)


#----------------------------------------------------
#create lagged variable
dv1 <- 
    impp2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1))
#code missings as 0
dv1[is.na(dv1)] <- 0

#calculate stability index
#subpentran, pension and ownwelfare are excluded
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 

#code irregular yb to 1.
yb[yb >1] <- 1 
mydata$stability_2 <- sqrt(yb)


#---------------------------------------------------------------------
#create lagged variable
dv1 <- 
    impp3 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1))
#code missings as 0
dv1[is.na(dv1)] <- 0

#calculate stability index
#subpentran, pension and ownwelfare are excluded
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 

#code irregular yb to 1.
yb[yb >1] <- 1 
mydata$stability_3 <- sqrt(yb)

#------------------------------------------------------------------------------
#create lagged variable
dv1 <- 
    impp4 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1))
#code missings as 0
dv1[is.na(dv1)] <- 0

#calculate stability index
#subpentran, pension and ownwelfare are excluded
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 

#code irregular yb to 1.
yb[yb >1] <- 1 
mydata$stability_4 <- sqrt(yb)
#---------------------------------------------------------------------------
#create lagged variable
dv1 <- 
    impp5 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_p = lag(expend_security_p, 1), 
           lag.expenddefence_p = lag(expenddefence_p, 1),
           lag.exp_public_order_p = lag(exp_public_order_p, 1), 
           lag.wagessalaries_p = lag(wagessalaries_p, 1), 
           lag.total_welfare_p = lag(total_welfare_p, 1), 
           lag.education_p = lag(education_p, 1), 
           lag.health_p = lag(health_p, 1),
           lag.social_protection_p = lag(social_protection_p, 1),
           lag.housing_p = lag(housing_p, 1))
#code missings as 0
dv1[is.na(dv1)] <- 0

#calculate stability index
#subpentran, pension and ownwelfare are excluded
attach(dv1)
yb <- (expend_security_p - lag.expend_security_p)^2 +
      (expenddefence_p - lag.expenddefence_p)^2 +
      (exp_public_order_p - lag.exp_public_order_p)^2 +
      (wagessalaries_p - lag.wagessalaries_p)^2 +
      (total_welfare_p - lag.total_welfare_p)^2 +
      (education_p - lag.education_p)^2 +
      (health_p - lag.health_p)^2 +
      (social_protection_p - lag.social_protection_p)^2 +
      (housing_p - lag.housing_p)^2 

#code irregular yb to 1.
yb[yb >1] <- 1 
mydata$stability_5 <- sqrt(yb)
```

```{r}
#recode 0 into missing 
mydata[, 90:95][mydata[, 90:95] == 0] <- NA

#output
write.csv(x=mydata, file = "mydata2.csv")
```

-->


