---
title: "LK-ordered-demo"
author: "Hao Wang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LK Scores for democracy and autocracies (3 categories)

```{r, message=FALSE, warning=FALSE}
library(Amelia)
library(dplyr)
library(Lmoments)
library(reshape2)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/mydata.csv")

#recode democracy index based on ordered scale
mydata$demo_order <- mydata$e_v2x_polyarchy_3C
mydata$demo_order[mydata$demo_order == 0] <- "auto"
mydata$demo_order[mydata$demo_order == 1] <- "comptauto"
mydata$demo_order[mydata$demo_order == 0.5] <- "democracy"

mydata$demo_order <- as.factor(mydata$demo_order)
```

```{r, message=FALSE, warning=FALSE}
imp0 <- select(mydata, cowcode, year, c(56:66), demo_order) 
#imp0 <- select(imp0)


#--------------------------------------------------------------------------
dv2 <- imp0
#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(diff.expend_security_EXP = expend_security_EXP - lag(expend_security_EXP, 1), 
           diff.expenddefence_EXP = expenddefence_EXP - lag(expenddefence_EXP, 1),
           diff.exp_public_order_EXP = exp_public_order_EXP - lag(exp_public_order_EXP, 1), 
           diff.wagessalaries_EXP = wagessalaries_EXP - lag(wagessalaries_EXP, 1), 
           diff.total_welfare_EXP = total_welfare_EXP - lag(total_welfare_EXP, 1), 
           diff.education_EXP = education_EXP - lag(education_EXP, 1), 
           diff.health_EXP = health_EXP - lag(health_EXP, 1),
           diff.social_protection_EXP = social_protection_EXP - lag(social_protection_EXP, 1),
           diff.housing_EXP = housing_EXP - lag(housing_EXP, 1), 
           diff.owelfarespend_EXP = owelfarespend_EXP - lag(owelfarespend_EXP, 1),
           diff.pensions_EXP = pensions_EXP - lag(pensions_EXP, 1))

dv2 <- select(dv2, demo_order, c(15:25))

aql <- melt(dv2, id.vars = c("demo_order"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$demo_order, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order <- t(b)[, 4]
lk_order
```

With GSRE data, LK score for ordered scale: autocracy is `r lk_order[1]`, competitive autocracy is `r lk_order[2]`, democracy is `r lk_order[3]`



# LK Score for always demcoracies/autocracies

??/stuck here unsolved


```{r, message=FALSE, warning=FALSE}
summary(mydata$demo_order)
#create an ID var identifying always demcoraccies and autocracies
#????????


```





# Exclude wage-salaries and pensions


```{r, message=FALSE, warning=FALSE}
dv3 <- select(dv2, -diff.wagessalaries_EXP, -diff.pensions_EXP)
aql <- melt(dv3, id.vars = c("demo_order"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$demo_order, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order <- t(b)[, 4]
lk_order


dv4 <- select(dv2, 
              -diff.wagessalaries_EXP, 
              -diff.pensions_EXP,
              -diff.expend_security_EXP,
              -diff.total_welfare_EXP)
aql <- melt(dv4, id.vars = c("demo_order"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$demo_order, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order4 <- t(b)[, 4]
lk_order4
```


Without wage and pension, LK score for ordered scale: autocracy is `r lk_order[1]`, competitive autocracy is `r lk_order[2]`, democracy is `r lk_order[3]`

Without wage, pension, total security spending and total welfare spending,  LK score for ordered scale: autocracy is `r lk_order4[1]`, competitive autocracy is `r lk_order4[2]`, democracy is `r lk_order4[3]`

The total security spending and total welfare spending should be excluded as they are calculated through the combination of other categories. Making them redundant. 




# Use var_GDPGSRE

Results show here. I need to run the code from the beginning to include the WDI index, 


```{r, message=FALSE, warning=FALSE}

imp0 <- select(mydata, cowcode, year, c(67:77), demo_order) 
#imp0 <- select(imp0)


#--------------------------------------------------------------------------
dv2 <- imp0
#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(diff.expend_security_GDPGSRE = expend_security_GDPGSRE - lag(expend_security_GDPGSRE, 1), 
           diff.expenddefence_GDPGSRE = expenddefence_GDPGSRE - lag(expenddefence_GDPGSRE, 1),
           diff.exp_public_order_GDPGSRE = exp_public_order_GDPGSRE - lag(exp_public_order_GDPGSRE, 1), 
           diff.wagessalaries_GDPGSRE = wagessalaries_GDPGSRE - lag(wagessalaries_GDPGSRE, 1), 
           diff.total_welfare_GDPGSRE = total_welfare_GDPGSRE - lag(total_welfare_GDPGSRE, 1), 
           diff.education_GDPGSRE = education_GDPGSRE - lag(education_GDPGSRE, 1), 
           diff.health_GDPGSRE = health_GDPGSRE - lag(health_GDPGSRE, 1),
           diff.social_protection_GDPGSRE = social_protection_GDPGSRE - lag(social_protection_GDPGSRE, 1),
           diff.housing_GDPGSRE = housing_GDPGSRE - lag(housing_GDPGSRE, 1), 
           diff.owelfarespend_GDPGSRE = owelfarespend_GDPGSRE - lag(owelfarespend_GDPGSRE, 1),
           diff.pensions_GDPGSRE = pensions_GDPGSRE - lag(pensions_GDPGSRE, 1))

dv2 <- select(dv2, demo_order, c(15:25))

aql <- melt(dv2, id.vars = c("demo_order"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$demo_order, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order <- t(b)[, 4]
lk_order



```


With _GDPGSRE variables, LK score for ordered scale: autocracy is `r lk_order[1]`, competitive autocracy is `r lk_order[2]`, democracy is `r lk_order[3]`


## Without wage pension


```{r, message=FALSE, warning=FALSE}
dv3 <- select(dv2, -diff.wagessalaries_GDPGSRE, -diff.pensions_GDPGSRE)
aql <- melt(dv3, id.vars = c("demo_order"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$demo_order, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order5 <- t(b)[, 4]
lk_order5


dv4 <- select(dv2, 
              -diff.wagessalaries_GDPGSRE, 
              -diff.pensions_GDPGSRE,
              -diff.expend_security_GDPGSRE,
              -diff.total_welfare_GDPGSRE)
aql <- melt(dv4, id.vars = c("demo_order"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$demo_order, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order6 <- t(b)[, 4]
lk_order6
```





Without wage and pension, LK score for ordered scale: autocracy is `r lk_order5[1]`, competitive autocracy is `r lk_order5[2]`, democracy is `r lk_order5[3]`

Without wage pension total security spending and total welfare spending, LK score for ordered scale: autocracy is `r lk_order6[1]`, competitive autocracy is `r lk_order6[2]`, democracy is `r lk_order6[3]`





# LK score for bivariate classification

```{r, warning=FALSE, message=FALSE}
mydata$bi_demo <- mydata$demo_order
mydata$bi_demo[mydata$bi_demo == "comptauto"] <- "auto"
mydata$bi_demo <- as.factor(mydata$bi_demo)

imp0 <- select(mydata, cowcode, year, c(56:66), bi_demo) 
#imp0 <- select(imp0)


#--------------------------------------------------------------------------
dv2 <- imp0
#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(diff.expend_security_EXP = expend_security_EXP - lag(expend_security_EXP, 1), 
           diff.expenddefence_EXP = expenddefence_EXP - lag(expenddefence_EXP, 1),
           diff.exp_public_order_EXP = exp_public_order_EXP - lag(exp_public_order_EXP, 1), 
           diff.wagessalaries_EXP = wagessalaries_EXP - lag(wagessalaries_EXP, 1), 
           diff.total_welfare_EXP = total_welfare_EXP - lag(total_welfare_EXP, 1), 
           diff.education_EXP = education_EXP - lag(education_EXP, 1), 
           diff.health_EXP = health_EXP - lag(health_EXP, 1),
           diff.social_protection_EXP = social_protection_EXP - lag(social_protection_EXP, 1),
           diff.housing_EXP = housing_EXP - lag(housing_EXP, 1), 
           diff.owelfarespend_EXP = owelfarespend_EXP - lag(owelfarespend_EXP, 1),
           diff.pensions_EXP = pensions_EXP - lag(pensions_EXP, 1))

dv2 <- select(dv2, bi_demo, c(15:25))

aql <- melt(dv2, id.vars = c("bi_demo"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$bi_demo, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
lk_order <- t(b)[, 4]
lk_order
```


With GSRE data, LK score for ordered scale: autocracy is `r lk_order[1]`, democracy is `r lk_order[2]`. 



# For Cross-Sectional OLS

## codes for generating LK score of each country

```{r, message=FALSE, warning=FALSE}
mydata$polity <- mydata$e_democ - mydata$e_autoc
  
#collapse original data to mean-substraction
d <- mydata %>%
 group_by(cowcode, country) %>%
 summarise_each(funs(mean(., na.rm = TRUE)))
#get rid of the missing
d <- d[!(is.na(d$stability_0)),]


imp0 <- select(mydata, cowcode, year, c(56:66), demo_order) 
#imp0 <- select(imp0)


#--------------------------------------------------------------------------
dv2 <- imp0
#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(diff.expend_security_EXP = expend_security_EXP - lag(expend_security_EXP, 1), 
           diff.expenddefence_EXP = expenddefence_EXP - lag(expenddefence_EXP, 1),
           diff.exp_public_order_EXP = exp_public_order_EXP - lag(exp_public_order_EXP, 1), 
           diff.wagessalaries_EXP = wagessalaries_EXP - lag(wagessalaries_EXP, 1), 
           diff.total_welfare_EXP = total_welfare_EXP - lag(total_welfare_EXP, 1), 
           diff.education_EXP = education_EXP - lag(education_EXP, 1), 
           diff.health_EXP = health_EXP - lag(health_EXP, 1),
           diff.social_protection_EXP = social_protection_EXP - lag(social_protection_EXP, 1),
           diff.housing_EXP = housing_EXP - lag(housing_EXP, 1), 
           diff.owelfarespend_EXP = owelfarespend_EXP - lag(owelfarespend_EXP, 1),
           diff.pensions_EXP = pensions_EXP - lag(pensions_EXP, 1))

dv2 <- select(dv2, cowcode, c(15:25))

aql <- melt(dv2, id.vars = c("cowcode"))
aql <- na.omit(aql)
a <- tapply(aql$value, aql$cowcode, function(x) Lcoefs(x))
b <- as.data.frame(matrix(unlist(a), nrow = 4))
#transpose, this is the LK score
d$lk_country <- t(b)[, 4]

```



## OLS

```{r, message=FALSE, warning=FALSE}
lm <- lm(lk_country ~ polity, data = d)
summary(lm)

#list cross-sectional country list
d2 <- select(d, country, lk_country, polity)
d2 <- na.omit(d2)
d2$country
```

## OLS with Quadratic Expression

```{r, message=FALSE, warning=FALSE}
qlm <- lm(lk_country ~ polity +I(polity^2), data = d)
summary(qlm)
```






------------------------------------------------------------------
# Number of Observations

```{r, message=FALSE, warning=FALSE}
library(Amelia)
library(dplyr)
library(Lmoments)
library(reshape2)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/mydata.csv")
GSRE <- select(mydata, c(56:66))
#if listwise deletion...
missmap(GSRE)
GSRE <- na.omit(GSRE)
nrow(GSRE)

#recode democracy index based on ordered scale
mydata$demo_order <- mydata$e_v2x_polyarchy_3C
mydata$demo_order[mydata$demo_order == 0] <- "auto"
mydata$demo_order[mydata$demo_order == 1] <- "comptauto"
mydata$demo_order[mydata$demo_order == 0.5] <- "democracy"

mydata$demo_order <- as.factor(mydata$demo_order)
```

Therefore, using list-wise deletion for GSRE part, only 87 cases remain. By cases, I mean country-year observations. Total observations are `r nrow(mydata)`. 


# How I calculated spendings

I use all categories of GSRE, as well as GDPGSRE, then I calculated wages and pensions and repeat the calculations. In calculating the LK score, I first calculate the between year differences of each country, then use the pooled difference items of each country to calculate the LK score of corresponding countries. 

In OLS regression, LK score is pooled score for each country, other independent variables are collapsed at mean levels. 




# Missingness and Political Regime

Most missing points are from the GSRE parts. 





```{r, echo=FALSE, warning=FALSE, message=FALSE}
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/mydata.csv")
GSRE <- select(mydata, c(56:66))

#recode democracy index based on ordered scale
mydata$demo_order <- mydata$e_v2x_polyarchy_3C
mydata$demo_order[mydata$demo_order == 0] <- "auto"
mydata$demo_order[mydata$demo_order == 1] <- "comptauto"
mydata$demo_order[mydata$demo_order == 0.5] <- "democracy"

mydata$demo_order <- as.factor(mydata$demo_order)

mydata$na.count <-  apply(GSRE, 1, function(x) sum(is.na(x)))
summary(lm(na.count ~ e_democ, data = mydata))
```






<!-- seems unrelevant for the moment

# Recalculating Stability and LK score

## add stability_x_aj to mydata.csv

```{r, message=FALSE}
library(dplyr)
library(Lmoments)
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/mydata.csv")

```




## without multiple imputation
```{r, message=FALSE}
dv2 <- select(mydata, c(56:66), cowcode, year)
#get rid of wage 
dv2 <- select(dv2, -pensions_EXP, -wagessalaries_EXP)


#replace NA to 0. 
dv2[is.na(dv2)] <- 0

#
#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_0_aj <- sqrt(yb)
#recode 0 to missing
mydata$stability_0_aj[mydata$stability_0_aj == 0] <- NA
```



# Multiple imputation: without wage index

```{r, message=FALSE, warning=FALSE}
library(dplyr)
imp1 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp1.csv")
imp2 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp2.csv")
imp3 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp3.csv")
imp4 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp4.csv")
imp5 <- read.csv("https://raw.githubusercontent.com/haowang666/Veto-Fiscal/master/data/imp5.csv") 

#-----------------------------------------------------------------------------------------------
dv2 <- imp1
#replace NA to 0. 
dv2[is.na(dv2)] <-0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_1_aj <- sqrt(yb)
#recode 0 to missing
mydata$stability_1_aj[mydata$stability_1_aj == 0] <- NA

#--------------------------------------------------------------------------
dv2 <- imp2
#replace NA to 0. 
dv2[is.na(dv2)] <- 0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_2_aj <- sqrt(yb)
#recode 0 to missing
mydata$stability_2_aj[mydata$stability_2_aj == 0] <- NA

#--------------------------------------------------------------------------
dv2 <- imp3
#replace NA to 0. 
dv2[is.na(dv2)] <- 0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_3_aj <- sqrt(yb)
#recode 0 to missing
mydata$stability_3_aj[mydata$stability_3_aj == 0] <- NA

#--------------------------------------------------------------------------
dv2 <- imp4
#replace NA to 0. 
dv2[is.na(dv2)] <- 0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_4_aj <- sqrt(yb)
#recode 0 to missing
mydata$stability_4_aj[mydata$stability_4_aj == 0] <- NA

#--------------------------------------------------------------------------
dv2 <- imp5
#replace NA to 0. 
dv2[is.na(dv2)] <- 0


#create lagged variable
dv2 <- 
    dv2 %>%
    group_by(cowcode) %>%
    mutate(lag.expend_security_EXP = lag(expend_security_EXP, 1), 
           lag.expenddefence_EXP = lag(expenddefence_EXP, 1),
           lag.exp_public_order_EXP = lag(exp_public_order_EXP, 1), 
           lag.total_welfare_EXP = lag(total_welfare_EXP, 1), 
           lag.education_EXP = lag(education_EXP, 1), 
           lag.health_EXP = lag(health_EXP, 1),
           lag.social_protection_EXP = lag(social_protection_EXP, 1),
           lag.housing_EXP = lag(housing_EXP, 1), 
           lag.owelfarespend_EXP = lag(owelfarespend_EXP, 1))

#calculate stability index
attach(dv2)
yb <- (expend_security_EXP - lag.expend_security_EXP)^2 +
     (expenddefence_EXP - lag.expenddefence_EXP)^2 +
     (exp_public_order_EXP - lag.exp_public_order_EXP)^2 +
     (total_welfare_EXP - lag.total_welfare_EXP)^2 +
     (education_EXP - lag.education_EXP)^2 +
     (health_EXP - lag.health_EXP)^2 +
     (social_protection_EXP - lag.social_protection_EXP)^2 +
     (housing_EXP - lag.housing_EXP)^2 +
     (owelfarespend_EXP - lag.owelfarespend_EXP)^2


#attach to data
mydata$stability_5_aj <- sqrt(yb)
#recode 0 to missing
mydata$stability_5_aj[mydata$stability_5_aj == 0] <- NA


#--------------------------------------------------------------------------
#export to doable file
write.csv(x = mydata, file = "mydata.csv", row.names = FALSE)
```


-->

